### 2020-09-28 tables/POPRES-matching.tex
code map

% latex table generated in R 3.6.1 by xtable 1.8-4 package
% Mon Sep 28 13:09:56 2020
\begin{table}[ht]
\centering
\caption{Ancestry (left) of POPRES individuals and their matching to 1000G populations (top) by our method. See the description of 1000G populations at \url{https://www.internationalgenome.org/category/population/}.}
\label{tab:ancestry-pred-popres}
\begin{tabular}{|l|c|c|c|c|c|c|}
  \hline
 & EUR\_CEU & EUR\_FIN & EUR\_GBR & EUR\_IBS & EUR\_TSI & NA \\
  \hline
Anglo-Irish Isles & 136 &  & 127 &  & 2 & 1 \\
  Belgium & 43 &  &  &  &  &  \\
  Central Europe & 47 &  &  &  & 8 &  \\
  Eastern Europe & 27 &  &  &  & 1 & 2 \\
  France & 49 &  & 3 & 35 & 2 &  \\
  Germany & 67 &  & 3 &  & 1 &  \\
  Italy & 1 &  &  & 11 & 204 & 3 \\
  Netherlands & 13 &  & 4 &  &  &  \\
  Scandinavia & 13 & 1 & 1 &  &  &  \\
  SE Europe & 12 &  &  & 3 & 70 & 9 \\
  SW Europe & 1 &  &  & 261 & 1 & 1 \\
  Switzerland & 179 &  &  & 32 & 11 &  \\
   \hline
\end{tabular}
\end{table}